---
title: "assign3"
author: "Hannah Cronin"
date: "2022-10-17"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(caret)
library(ISLR)
library(e1071)
library(readr)
library(ggplot2)
library(lattice)
library(reshape)
library(reshape2)
UniversalBank <- read_csv("Desktop/GITHUB/64060_-HCRONIN-FML/UniversalBank.csv")
```

```{r}
loan = UniversalBank[,-c(1,2,3,4,5,6,7,8,9,11,12)] #more narrowed dataset
colnames(loan)[1] = 'Personal_Loan'
```

```{r}
Index_Train = createDataPartition(loan$`Personal_Loan`, p = .6, list=FALSE)
Train = loan[Index_Train,]
Valid = loan[-Index_Train,]
```

A.
```{r}
pivot = melt(Train, id = c('CreditCard','Personal_Loan'), variable = 'Online')
cast.loan = cast(pivot, CreditCard+Personal_Loan~Online)
cast.loan[, c(1:3)]
```
B.
```{r}
prob = 87 / (1919+206+788+87)
prob * 100
# The probability that this customer will accept a loan is 2.90%.
```

C.
```{r}
pivot2 = melt(Train, id = c('Personal_Loan'), variable = 'Online')
```
```{r}
cast.loan2 = cast(pivot2, Personal_Loan~Online)
cast.loan2[,c(1:2)]
```

```{r}
pivot3 = melt(Train, id = c('Personal_Loan'), variable = 'CreditCard')
```
```{r}
cast.loan3 = cast(pivot3, Personal_Loan~CreditCard)
cast.loan3[,c(1,3)]
```
# I can't get the pivot tables above to display the correct answer - they seem to be defaulted to the number of people who accepted the loan regardless of the other variable we're looking at. (293 & 2707)
# I'm going to use regular tables instead so at least I can compute the right answers for D.
```{r}
table(Train[,c(1,2)])
table(Train[,c(1,3)])
```

D.
```{r}
# 1. 29.693%
((87)/(206+87))
# 2. 60.751%
(178/(115+178))
# 3. 9.767%
(293/3000)
# 4. 29.110 %
(788/(788+1919))
# 5. 59.660
(1615/(1615+1092))
# 6. 90.233%
((3000-293)/3000)
```

E.
```{r}
((87/(206+87))*(178/(178+115))*(293/(293+2707)))/(((77/(77+198))*(178/(178+115))*(293/(293+2707)))+((788/(788+1919))*(1615/(1615+1092))*2707/(2707+293)))
```
F.
10.16% and 10.73% are very similar. Since Naive assumes independence of the variables- it's a more accurate estimate.

G.
```{r}
nb_model <-naiveBayes(Personal_Loan~CreditCard+Online,data = Train)
nb_model
(.458)*(.489)*(.10)/((.458*.489*.098)+(.454*.491*.90))
```
This number is very close to the one found in E.
