---
title: "Assignment_4"
author: "Hannah Cronin"
date: "2022-10-30"
output: html_document
---
```{r}
library(readr)
library(tidyverse)
library(factoextra)
library(ISLR)
library(cluster)
Pharmaceuticals <- read_csv("/Users/hannahcronin/Desktop/GITHUB/64060_-HCRONIN-FML/Assignment_4/Pharmaceuticals.csv")
```

```{r}
set.seed(123)
df = Pharmaceuticals[, c(3,4,5,6,7,8,9,10,11)]
rownames(df) <- c('ABT','AGN','AHM','AZN','AVE','BAY','BMY','CHTT','ELN','LLY','GSK','IVX','JNJ','MRX','MRK','NVS','PFE','PHA','SGP','WPI','WYE')
colnames(df) <- c('Market Cap','Beta','PE_Ratio','ROE','ROA','Asset_Turnover','Leverage','Rev_Growth','Net_Profit_Margin')
summary(df)
```

```{r}
df = scale(df) #to normalize data
rownames(df) <- c('ABT','AGN','AHM','AZN','AVE','BAY','BMY','CHTT','ELN','LLY','GSK','IVX','JNJ','MRX','MRK','NVS','PFE','PHA','SGP','WPI','WYE') #row names kept disappearing
colnames(df) <- c('Market Cap','Beta','PE_Ratio','ROE','ROA','Asset_Turnover','Leverage','Rev_Growth','Net_Profit_Margin') #also to ensure column names stick
distance = get_dist(df)
fviz_dist(distance)
summary(df)
```
I used the silhouette method to find the best number of clusters.
```{r}
fviz_nbclust(df, kmeans, method = "silhouette")
```

Running K-Means
```{r}
km = kmeans(df, 5, nstart = 25)
km$cluster
```
I decided to use KMeans/Euclidean distance because these financial ratios/statistics are not inherently correlated. A few may share some common demoninators, however they represent/pull other data from such different areas (ex: different financial statements) that I chose not to use the Manhattan distance metric.

Cluster 1: ABT, AHM, AZN, BMY, LLY, NVS, SGP, WYE
Cluster 2: BAY, CHTT, IVX, 
Cluster 3: AGN, WPI
Cluster 4: GSK, JNJ, MRK, PFE
Cluster 5: AVE, ELN, MRX, WPI
```{r}
plot(c(0), xaxt = 'n', ylab = "", type = "l", 
     ylim = c(min(km$centers), max(km$centers)), xlim = c(0, 9))
axis(1, at = c(1:9), labels = names(df), cex.axis=.8)
for (i in c(1:5))
lines(km$centers[i,], lty = i, lwd = 2, col = ifelse(i %in% c(1, 3, 5),
                            "blue", "red"))
text(x = 0.3, y = km$centers[, 1], labels = paste("Cluster", c(1:5)))

```

Descriptions of each cluster:
Cluster 1 = Low Beta, Low Rev_Growth, High Net_Profit_Margin (No extremes)
Cluster 2 = Low Market_Cap, High Beta, Low ROA, High leverage
Cluster 3 = High PE_Ratio, Low Net_Profit_Margin
Cluster 4 = High Market_Cap, High ROE, High ROA, High Asset_Turnover, High Net_Profit_Margin
Cluster 5 = Low PE_Ratio, Low Asset_Turnover
```{r}
km$centers #numerical descriptions of each cluster
km$withinss #numerical descriptions of each cluster
km$size #numerical descriptions of each cluster
```
Clusters 3,4 are the most homogenous- also amongst the smallest clusters. The least homogenous is Cluster 1, coincidentally also the largest cluster.
```{r}
dist(km$centers) 
```
```{r}
fviz_cluster(km, data = df) # Visualize the output
```

Patterns among the clusters (categorical variables):
Cluster 1: All traded on the NYSE, 5/8 of these are US companies
Cluster 2: 2/3 of these are holds, 2/3 are US companies
Cluster 3: Both listed on the NYSE
Cluster 4: 2 are moderately buy, 2 are hold, 3/4 are US companies, all traded on NYSE
Cluster 5: 2 Moderately sell, 2 moderately buy, all traded on NYSE

When it comes to the categorical variables, there's definite trends in the dataset related to trading environment and location- however since US and NYSE dominate both categories, I don't think these similiarities are related to the clustering. From the clusters that my model generated, there were no overhwhelming similarities when it came to the median_recommmendation for each company.

Names for each cluster:
Cluster1: The_Middle_No_Extremes
Cluster2: Low_Market_Cap_ROA_High_Beta_Leverage
Cluster3: High_PE_Ratio_Low_Net_Profit_Margin
Cluster4: High_Market_Cap_ROE_ROA_Asset_Turnover_Net_Profit_Margin
Cluster5: Low_PE_Ratio_Asset_Turnover

